### 2017-05-01 project/proposal/execsummary.tex

\chapter*{Executive Summary}
The Office of Strategic National Alien Planning (OSNAP) has made a substantial investment in a Secure Alien Surveillance System (SASS). As part of the acquisition, control software is required to track potential targets in surveillance video using Kalman Filters.

To achieve the necessary performance while tracking multiple objects the Kalman Filters will be run in parallel with each other and additionally use parallelism to improve the performance of the underlying calculations. The project will produce a demo system that incorporates multiple types of parallelism into a tracking system to identify and track objects as they move through recorded video.